\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-09-26 distributed-estimation.tex
Fixed proof of (previously) Lemma 15 and made changes in structure regarding this lemma. Now one lemma (Lemma 12) is used in three different results.

\begin{thm} \label{thm:LAN1}
For $n \in \mathbb N$ and $A_n \subset \mathbb R$,  define
\begin{equation}
\label{eq:precision_general}
L_n(A_1,\ldots,A_n;\theta) \triangleq \frac{1}{n} \sum_{i=1}^n \frac{ \left(\frac{d}{d \theta} \Prob(X_i \in A_i) \right)^2 }{ \Prob(X_i \in A_i)\left(1- \Prob(X_i \in A_i) \right) }. 
\end{equation}
Consider the following conditions:
\begin{itemize}
\item [(i)\,\,] The pdf $f(x)$ of $X_n-\theta$ is a log-concave, differentiable and symmetric density function such that $\eta(x)$ is unimodal.
\item[(ii)\,] $A_n$ is a finite union of disjoint intervals.
\item[(iii)] The limit 
\begin{equation}
\label{eq:LAN_lim}
\kappa(\theta) \triangleq \lim_{n\to \infty} L_n(A_1,\ldots,A_n; \theta)
\end{equation}
exists. 
\end{itemize}
For $i=1,\ldots,n$ set
\[
B_n = \begin{cases} 1 & X_n \in A_n, \\
-1 & X_n \notin A_n. 
\end{cases} 
\]
For any $\theta$, $f(x)$ and a sequence of sets $A_1,A_2,\ldots$ such that (i)-(iii) hold, and any $h\in \mathbb R$, we have
\begin{align*}
& \log  \frac{ \mathbb P_{\theta+h/\sqrt{n}} (B_1,\ldots,B_n) }{
\mathbb P_{\theta} (B_1,\ldots,B_n)} \\
& \overset{d}{\longrightarrow} \Ncal\left(-\frac{1}{2} h^2 \kappa(\theta), h^2 \kappa(\theta) \right).
\end{align*} 
\end{thm}

@@

As can be seen in Fig.~\ref{fig:opt_density} for the case $P_X = \Ncal(\theta ,\sigma^2)$ and a uniform $\pi$, when the radius of $\Theta$ is small compared to $\sigma$, the optimal distribution $\lambda^\star$ is a mass distribution. In this case, the ML estimator reduces to the estimator ${\theta}_n$ of \eqref{eq:estimator_naive}. As the following proposition shows, the Bayes risk for this choice of $\lambda$ is maximal, and thus provides an upper bound on the Bayes risk under any $\lambda$. 
 \eqref{eq:cvx_average}. 
\begin{prop}\label{prop:upper_bound}
For any prior $\pi(d\theta)$ and $\theta_0$ in the support of $\eta(x)$ we have
\begin{equation} 
\label{eq:upper_bound}
R_\pi^\star  \leq 
 \int \frac{\pi(d\theta)}{\eta \left( \theta_0 - \theta \right)},
\end{equation}
Furthermore, assuming that $\theta_0 = \mathbb E \theta$ and that $\pi$ has a finite third moment $\sigma_\theta^3$, we have: 
\begin{equation}
\label{eq:bound_Taylor}
R^\star_\pi \leq \frac{1}{4 f^2(0)} + \left(\frac{1}{4 f^2(0)} \frac{-f''(0)}{f(0)} -1 \right) \sigma_\theta^2 + O(\sigma_\theta^3).
\end{equation}
\end{prop}
\begin{proof}
The function $x \rightarrow 1/x$ is convex for positive values, hence Jensen's inequality implies
\[
\left( \int \eta \left( t-\theta\right) \lambda(dt) \right)^{-1}  \leq \int  \frac{ \lambda(dt)}{ \eta \left( t-\theta\right)  }. 
\]
Therefore, the expected value of $\kappa^{-1}(\theta)$ satisfies
\begin{align}
\ex{  \frac{1}{\kappa(\theta)}}  \leq \int \int \frac{\pi(d\theta) \lambda(dt) }{\eta \left( t - \theta \right)}. \label{eq:upper_bound_proof}
\end{align}
The bound \eqref{eq:upper_bound} is obtained by taking $\lambda$ to be a mass distribution at any $\theta_0$ in the support of $\eta(x)$. Finally, \eqref{eq:bound_Taylor} is obtained by expanding $1/\eta(x)$ to a third order Taylor series around zero
%\[
%1/\eta(x) = \frac{1}{4f^2(0)}  + \left(\frac{1}{4 f^2(0)} \frac{-f''(0)}{f(0)} -1 \right)  x^2 + O(x^3),
%\]
and taking its expectation with respect to $\pi$ at $x=\theta_0-\theta$. 
\end{proof}
We note that the function $1/\eta(x)$ is quasi-convex and symmetric around zero, so the choice $\theta_0 = \ex{ \theta}$ minimizes the RHS of \eqref{eq:upper_bound} among all $\theta_0$ in the support of $\eta(x)$.\par
The bound \eqref{eq:upper_bound} is not trivial as long as the integral in the RHS of \eqref{eq:upper_bound} is finite, i.e., whenever the tail of $\pi(\theta)$ vanishes fast enough compared to $\eta(x)$. The expansion \eqref{eq:bound_Taylor} implies that this bound becomes tight whenever the support of the optimal distribution is a mass distribution at $\mathbb E \theta$, in which case the expected value of $\kappa^{-1}(\theta)$ approaches $1/\eta(0) = 1/4f^2(0)$.

>>>>

\begin{thm} \label{thm:LAN1}
For $n \in \mathbb N$ and $A_n \subset \mathbb R$,  define
\begin{equation}
\label{eq:precision_general}
L_n(A_1,\ldots,A_n;\theta) \triangleq \frac{1}{n} \sum_{i=1}^n \frac{ \left(\frac{d}{d \theta} \Prob(X_i \in A_i) \right)^2 }{ \Prob(X_i \in A_i)\left(1- \Prob(X_i \in A_i) \right) }. 
\end{equation}
Consider the following conditions:
\begin{itemize}
\item [(i)\,\,] The pdf $f(x)$ of $X_n-\theta$ is log-concave, differentiable, symmetric,  and  there exists $\delta>0$ such that $\eta^{1+\delta}(x)/f^\delta(x)$ has a unique maximum at the origin.
\item[(ii)\,] $A_n$ is a finite union of disjoint intervals.
\item[(iii)] The limit 
\begin{equation}
\label{eq:LAN_lim}
\kappa(\theta) \triangleq \lim_{n\to \infty} L_n(A_1,\ldots,A_n; \theta)
\end{equation}
exists. 
\end{itemize}
For $i=1,\ldots,n$ set
\[
B_n = \begin{cases} 1 & X_n \in A_n, \\
-1 & X_n \notin A_n. 
\end{cases} 
\]
For any $\theta$, $f(x)$ and a sequence of sets $A_1,A_2,\ldots$ such that (i)-(iii) hold, and any $h\in \mathbb R$, we have
\begin{align*}
& \log  \frac{ \mathbb P_{\theta+h/\sqrt{n}} (B_1,\ldots,B_n) }{
\mathbb P_{\theta} (B_1,\ldots,B_n)} \\
& \overset{d}{\longrightarrow} \Ncal\left(-\frac{1}{2} h^2 \kappa(\theta), h^2 \kappa(\theta) \right).
\end{align*} 
\end{thm}

@@

As can be seen in Fig.~\ref{fig:opt_density} for the case $P_X = \Ncal(\theta ,\sigma^2)$ and a uniform $\pi$, when the radius of $\Theta$ is small compared to $\sigma$, the optimal distribution $\lambda^\star$ is a mass distribution. In this case, the ML estimator reduces to the estimator ${\theta}_n$ of \eqref{eq:estimator_naive}. As the following proposition shows, the Bayes risk for this choice of $\lambda$ is maximal, and thus provides an upper bound on the Bayes risk under any $\lambda$. 
 \eqref{eq:cvx_average}. 
\begin{prop}\label{prop:upper_bound}
For any prior $\pi(d\theta)$ and $\theta_0$ in the support of $\eta(x)$ we have
\begin{equation} 
\label{eq:upper_bound}
R_\pi^\star  \leq 
 \int \frac{\pi(d\theta)}{\eta \left( \theta_0 - \theta \right)},
\end{equation}
Furthermore, assuming that $\theta_0 = \ex{\theta}$ and that $\pi$ has a finite third moment $\sigma_\theta^3$, we have: 
\begin{equation}
\label{eq:bound_Taylor}
R^\star_\pi \leq \frac{1}{4 f^2(0)} + \left(\frac{1}{4 f^2(0)} \frac{-f''(0)}{f(0)} -1 \right) \sigma_\theta^2 + O(\sigma_\theta^3).
\end{equation}
\end{prop}
\begin{proof}
The function $x \rightarrow 1/x$ is convex for positive values, hence Jensen's inequality implies
\[
\left( \int \eta \left( t-\theta\right) \lambda(dt) \right)^{-1}  \leq \int  \frac{ \lambda(dt)}{ \eta \left( t-\theta\right)  }. 
\]
Therefore, the expected value of $\kappa^{-1}(\theta)$ satisfies
\begin{align}
\ex{  \frac{1}{\kappa(\theta)}}  \leq \int \int \frac{\pi(d\theta) \lambda(dt) }{\eta \left( t - \theta \right)}. \label{eq:upper_bound_proof}
\end{align}
The bound \eqref{eq:upper_bound} is obtained by taking $\lambda$ to be a mass distribution at any $\theta_0$ in the support of $\eta(x)$. Finally, \eqref{eq:bound_Taylor} is obtained by expanding $1/\eta(x)$ to a third order Taylor series around zero
%\[
%1/\eta(x) = \frac{1}{4f^2(0)}  + \left(\frac{1}{4 f^2(0)} \frac{-f''(0)}{f(0)} -1 \right)  x^2 + O(x^3),
%\]
and taking its expectation with respect to $\pi$ at $x=\theta_0-\theta$. 
\end{proof}
We note that the function $1/\eta(x)$ is quasi-convex and symmetric around zero, so the choice $\theta_0 = \ex{ \theta}$ minimizes the RHS of \eqref{eq:upper_bound} among all $\theta_0$ in the support of $\eta(x)$.\par
The bound \eqref{eq:upper_bound} is not trivial as long as the integral in the RHS of \eqref{eq:upper_bound} is finite, i.e., whenever the tail of $\pi(\theta)$ vanishes fast enough compared to $\eta(x)$. The expansion \eqref{eq:bound_Taylor} implies that this bound becomes tight whenever the support of the optimal distribution is a mass distribution at $\mathbb E \theta$, in which case the expected value of $\kappa^{-1}(\theta)$ approaches $1/\eta(0) = 1/4f^2(0)$.
